---
title: "Seurat_Code_Filtering_2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Install packages
```{r}
#install.packages("BiocManager")
#BiocManager::install("Seurat")
#BiocManager::install("ggplot2")
#install.packages("dplr")
#install.packages("ape")
#install.packages("cowplot")
#install.packages("Matrix")
#install.packages("clustree")
#install.packages("igraph")
install.packages('patchwork')
install.packages('stringr')
BiocManager::install('Nebulosa')
install.packages("qs")
# Installing scTransform v2
# install glmGamPoi
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::install("glmGamPoi")

# install sctransform from Github
install.packages("devtools")
library(devtools)
devtools::install_github("satijalab/sctransform", ref = "develop")
devtools::install_github(repo = "samuel-marsh/scCustomize")

if (!require("devtools"))
  install.packages("devtools")
devtools::install_github("Winnie09/Palo")

BiocManager::install("glmGamPoi")
install.packages('Hmisc')
```

Load packages into your workspace. 
```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(ape)
library(cowplot)
library(Matrix)
library(sctransform)
library(clustree)
library(tidyverse)
library(magrittr)
library(RColorBrewer)
library(patchwork)
library(glmGamPoi)
library(scCustomize)
library(qs)
library(Nebulosa)
library(Palo)
library(Hmisc)
```


```{r}
#set up folders
QCfolder = "QC"
Plotfolder = "Plots"

#Specify your organism
organism = "Mouse"

#metrics you want to look at for QC
m = c("nCount_RNA","nFeature_RNA","percent.mito")
```


Load in your filtered dataset
```{r}
dir= "/Users/phinguyen/Library/CloudStorage/Dropbox/Hen Lab/Data/SingleCell RNA-seq/2_AllCells_Early-Late-Flx_June 2023/Seurat/Run 1/Raw/2_all cells_afterQC_Run 1.RData"

load(file.path(dir))
```

Function to print multiple graphs: 
```{r}
PrintSeuratGraph = function(namecard = "a",seurat_object = sobject,graphtype = "feature",feature = NULL,group = NULL,split=NULL,cellnames=NULL){
  if (!is.null(cellnames)){
    Idents(seurat_object) = cellnames[1]
    cells = colnames(seurat_object)[Idents(seurat_object) %in% cellnames[2:length(cellnames)]]} 
  else {cells = cellnames}
  if (graphtype == "feature"){
    graph = FeaturePlot(seurat_object,features = feature,split.by = split, cells = cells,cols = c("lightyellow","darkred"))
  }
  if (graphtype == "violin"){
    graph = VlnPlot(seurat_object,features = feature, pt.size = 0.1, idents = cellnames[2:length(cellnames)],group.by = group, split.by = split)
  }
  if (graphtype == "dim"){
    graph = DimPlot(seurat_object,cells = cells, group.by = group, split.by = split)
    
  }
  
  name = paste0(feature,"_",graphtype,namecard,".eps")
  graph
  setEPS()
  postscript(file.path(dir,Plotfolder,name))
  print(graph) 
  dev.off()
}

```



*##############################**##############################*

```{r}
#options(future.globals.maxSize = 4000 * 1024^5)
options(future.globals.maxSize = 8000 * 1024^5)

library(usethis) 
usethis::edit_r_environ()
```


#Find variable features, normalize, scale, run PCA, clustering, umap
Normalization and variance stabilization (Hafemeister and Satija, 2019 https://pubmed.ncbi.nlm.nih.gov/31870423/)
```{r, results='hide',eval = FALSE}
regress = c("nCount_RNA", "percent.mito") #"orig.ident"

sobject <- SCTransform(sobject, vars.to.regress = regress, assay = "RNA", verbose = T, variable.features.n = 3000, 
                       conserve.memory = T, vst.flavor = NULL) 

# can set ncells = 5000 to speed up

                         #vst.flavor = "v2" #vst.flavor v2 specifies SCtransform version 2
                        # method = "glmGamPoi"
# "limit reachediteration" error message is ok. Good idea to save prior to this step b/c may crash R.
```

Show most variable genes
```{r}
labels <- c(head(VariableFeatures(sobject),20))
plot1 = VariableFeaturePlot(sobject)
LabelPoints(plot=plot1, points = labels, repel = F, xnudge = 0.1, ynudge = 0.5)
```

*##############################**##############################*

*Run PCA analysis and show elbow plot*
```{r}
sobject <- RunPCA(sobject,features = VariableFeatures(sobject),npcs = 50, verbose = FALSE)
ElbowPlot(sobject,ndims = 50, reduction = "pca")


print(sobject[["pca"]], dims = 1:20, nfeatures = 15)
# Output will be 5 genes per PC. PCs enriched with Mito/heatshock/ribo genes can be excluded

# Plot PCA
PCAPlot(sobject, split.by = "condition", pt.size = 1)  
PCAPlot(sobject, group.by = "condition", pt.size = 1)  

#scCustomize heatmap of defined PC
PC_Plotting(seurat_object = sobject, dim_number = 1)
```


Which PC's to include for downstream clustering? Can estimate qualitatively based on elbow plot or take quantitative approach to find PC's in which: 
1.) The point where the PCs only contribute 5% of standard deviation and the principal components cumulatively contribute 90% of the standard deviation.
2.) The point where the percent change in variation between the consecutive PCs is less than 0.1%.
```{r}
# Find the first metric (1)
# Determine percent of variation associated with each PC
pct <- sobject[["pca"]]@stdev / sum(sobject[["pca"]]@stdev) * 100

# Calculate cumulative percents for each PC
cumu <- cumsum(pct)

# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
metric1 <- which(cumu > 90 & pct < 5)[1]
print("Last PC where cumulative % of SD is 90%")
metric1

# Find the second metric (2)
# Determine the difference between variation of PC and subsequent PC
metric2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1
print("Last PC where % change of variation is more than 0.1%.")
metric2
```

```{r}
# PCs for dimensionality reduction. Can change if one of the pc's is for something you think is a technical error (i.e. HSP, RP, etc)
pcs = c(1:35)

print(sobject[["pca"]], dims = 1:20, nfeatures = 8)
```

Once you are satisfied with pc's, run clustering (UMAP, FindNeighbors, FindClusters): 
```{r}
# number of PC's defined at top of chunk above
sobject<-RunUMAP(sobject,reduction = "pca",dims = pcs, verbose = F) 


# Determine the K-nearest neighbor graph
sobject<-FindNeighbors(sobject,dims=pcs,verbose=F) 

# Quick UMAP test
Idents(sobject) = "condition" 

pt.size = 0.1

DimPlot(sobject, label = F, label.size = 4, pt.size = pt.size, 
        label.color = "black",repel = F,
        raster = F) +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 
```


Determine the clusters for various resolutions
```{r}
# clustering resolution; the last number will be saved as "seurat_clusters" in metadata
res = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6)
sobject<-FindClusters(sobject,verbose=F,resolution = res)
# sobject<-FindClusters(sobject,verbose=F,resolution = 0.1)


#Test how UMAP plot looks with different clustering resolutions
Idents(sobject) = "seurat_clusters"

for (Cluster_resolution in c("SCT_snn_res.0.1",
                             "SCT_snn_res.0.2",
                             "SCT_snn_res.0.3",
                             "SCT_snn_res.0.4",
                             "SCT_snn_res.0.5",
                             "SCT_snn_res.0.6")){
  Idents(sobject) = Cluster_resolution
  print(DimPlot(sobject, label = T, label.size = 6, pt.size = 1.0) + ggplot2::ggtitle(label = Cluster_resolution) +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 
        )
      }
```


Clustree package to generate clustering trees- see how cells/clusters move as clustering resolution increases. Vignette: https://cran.r-project.org/web/packages/clustree/vignettes/clustree.html 
```{r}
clustree(sobject, prefix = "SCT_snn_res.") + theme(legend.position = "bottom")


# Overlay expression of gene of interest

#clustree(sobject, prefix = "SCT_snn_res.", layout = "sugiyama", use_core_edges = FALSE, node_colour = "Prox1", node_colour_aggr = "mean") + guides(edge_colour = FALSE, edge_alpha = FALSE) + theme(legend.position = "bottom") + scale_color_gradient(low = "lightblue", high = "red")
    
#clustree(sobject, prefix = "SCT_snn_res.", 
        # node_colour = "nFeature_SCT", node_colour_aggr = "mean", 
        # node_label = "nFeature_SCT", node_label_aggr = "max")

#clustree(sobject, prefix = "SCT_snn_res.", 
        # node_colour = "sc3_stability") #higher sc3_stability, the more stable the cluster across resolutions

# Opacity of arrows indicates relative contribution of cells moving towards a cluster of a subsequent clustering resolution
```


*##############################**##############################*


```{r}
# save

dir = "/Users/phinguyen/Library/CloudStorage/Dropbox/Hen Lab/Data/SingleCell RNA-seq/2_AllCells_Early-Late-Flx_June 2023/CellRanger-6.1.2/Seurat/Run 2"

save(sobject,file = file.path(dir,paste0("2_all cells_2024-8-13.RData")))
```

