---
title: "Gene lists"
output: html_document
date: "2024-07-04"
---


```{r}
library(Hmisc)
library(scCustomize)
library(openxlsx)
library(ggdark)
library(Seurat)
library(ggplot2)
```


Top 100 genes from From Hochgerner_age_celltype_P132_GC-adult vs P5_Immature-GC
Hochgerner et al., 2018, Nat Neuro
*P5_GC_1*
*P132_GC_1*
```{r}
# From Hochgerner_age_celltype_P132_GC-adult vs P5_Immature-GC

# Top 100 Genes that are upregulated in juvenile GCs (defined by FindMarkers comparison of P132_GC-adult vs P5_Immature-GC metadata clusters)
P5_GC <- Cs(
Crip2,
Pgls,
Cdk4,
C530008M17Rik,
Mapk6,
L1cam,
Prkra,
Neurod1,
Zfp57,
Cadm4,
Csdc2,
Atat1,
Sbk1,
Sh3yl1,
Sqle,
Arhgdig,
Marcksl1,
Camk4,
Pgap1,
Fgfr1,
Elavl2,
Tubb5,
Ier2,
H2afy2,
Rerg,
Mad2l2,
Kif21b,
Gm9844,
Syce2,
Stmn1,
Nans,
Fabp7,
Fxyd6,
Gm11837,
Gm17750,
Rnd2,
Ppp1r14c,
Myl6b,
Idh1,
Rac3,
Elmo1,
Dner,
Eps8,
Rnf121,
Cntnap2,
Gm10320,
Tmsb10,
Sox11,
Fabp5,
Pafah1b3,
Tuba1a,
Csrp2,
Cks1b,
Pcdh15,
Dap,
Sox4,
Cxadr,
Tubb3,
Shd,
Adamts18,
Midn,
Sstr2,
Meis2,
Rgs6,
Mtss1,
Smim18,
Gap43,
Dcx,
Stmn2,
Hmgb3,
Nnat,
Dpysl5,
Frmd3,
Fezf2,
Traf4,
Sv2b,
Arl4c,
Bach2,
Carhsp1,
Ttc28,
Tubb2b,
Dpysl3,
Rbp1,
Fos,
Socs2,
Neurod6,
Calb2,
Ddah2,
Hba-a1,
Fxyd7,
Col12a1,
Mex3a,
Fnbp1l,
Trim67,
St8sia2,
Hbb-bs,
Sema3c,
Cd24a,
Draxin,
Crabp1,
Igfbpl1,
)

# Top 100 Genes that are up in mature GCs (defined by FindMarkers comparison of P132_GC-adult vs P5_Immature-GC metadata clusters)
P132_GC <- Cs(
Rassf8,
Hlf,
Ano3,
Zmym1,
Nptx1,
Tanc1,
Rasl10a,
Gng13,
Scn2b,
Zfp365,
Slc30a3,
Ptk2b,
Ddn,
Gas7,
Fam19a1,
Ahcyl2,
Atp1a3,
Pitpnm2,
PISD,
Syne1,
Cpne6,
Grin2a,
Ncdn,
Galnt1,
Vstm2a,
Faim2,
Ntng1,
Dock10,
Mt1,
Mkl2,
Gabra1,
Slc4a10,
Camkk1,
Slc24a2,
Camk2n1,
Klhl2,
Nrsn1,
Sorl1,
Lypd1,
Gm10605,
Mink1,
Cdk17,
Itm2c,
A830018L16Rik,
Chn1,
Tmem243,
Slc12a5,
Lgi1,
Wipf3,
Ppp3r1,
Tppp,
Unc80,
Map2k1,
Pcdh20,
Agap2,
Rprml,
Dgkz,
Ndrg4,
Gm26825,
Hpca,
Dnm1,
Hpcal4,
Nmt1,
Ptprn,
Cox17,
Rasgrf1,
Gabra4,
Lmo3,
Phyhip,
Tef,
Cdk14,
Pcdh19,
Slc7a14,
Wasf1,
Pgbd5,
Ppp3ca,
Adcy1,
Nrn1,
Eno2,
Scn1b,
Camk2a,
Abr,
Fam126b,
Col25a1,
Pcsk2,
Kdm7a,
Slc17a7,
Gria3,
Reps2,
Pgm2l1,
Atp6v1g2,
Pop5,
Gm27032,
Kcnj3,
Rapgef2,
Mapk1,
Ramp1,
App,
Wbp11,
Cnksr2,
Impact
)

sobject <- AddModuleScore(sobject,
                  features = list(P5_GC),
                  name= "P5_GC_")

sobject <- AddModuleScore(sobject,
                  features = list(P132_GC),
                  name= "P132_GC_")
```

Regeneration-Associated Genes 
Matson et al., 2022, Nature Communications
*RAG_1*
```{r}
RAG <- Cs(Atf3, Sprr1a, Tnfrsf12a, Sox11, Klf6, Bdnf, Adcyap1, Gap43)

sobject <- AddModuleScore(sobject,
                  features = list(RAG),
                  name= "RAG_")
```

4 week abGC vs 8 week abGC - ascl-creER x sun1-GFP
Rasetto et al 2024 Sci Adv from Alejandro Schinder
*Four_wk_abGC_1*
```{r}
#top 100 genes upregulated in 4w abGCs vs 8w

Four_wk_abGC <- Cs(Cacng4, Gm15398, Gm26905, Kcnip1,Fam163a, Gm16159, Lemd1, Col15a1, Socs2, Triobp, Macrod1, Aldh1l2, Draxin,Gm26936, Fgf1, Filip1l, Mad2l2, Nkain1, Adamts18, Myo16, Taco1, Shb, Gm26625, Nell1, Vav2, B430010I23Rik, Sema4g, Tmem44, Lypd6, Sox11, Slc17a6, Chn2, A930011G23Rik, Sox4, Kcnh8, Gm11146, Gm16279, Pard3bos1, Zfp57, Flrt1, Clcn5, Zfp558, Dner, Fnbp1l, Cntnap4, Gmip, Galntl6.1, Gm4128, Phykpl, Pcdh15, Adam12, Elavl2, Sh3bp2, Lppr3, Gm13944, Igdcc3, Sstr2, Sox2ot, Dpysl3, Adam19, Gm15645, Col11a1, Gm15563, Fam196a, St6gal1, C530008M17Rik, Mtss1, Col12a1, Sh3bp4, Calb2, Igfbpl1, Olfm2, Slc24a5, Gm26735, Fhod3, Pappa2, Psd2, Slco5a1, Myo1b, L1cam, St8sia2, Prrg1, Ccdc88c, Angpt1, Podxl2, Gm37679, Cnr1, Kif21b, Bahcc1, Prkd3, Gm36992, Hunk, Rgs6, Syt9, Frmd3)

sobject <- AddModuleScore(sobject,
                  features = list(Four_wk_abGC),
                  name= "Four_wk_abGC_")
```

DG GCs after 2mo EE (young mice) vs home cage
from Raul et al., 2024 Nat Comms (https://pubmed.ncbi.nlm.nih.gov/39013876/)
*EE_top500_1*
*EE_top100_1*
```{r}
EE_top500 <- Cs(Nrp1, AC149090.1, Pcdh7, Nrxn3, Tnr, C1ql3, Chst9, Malat1, Hspa5, Dgkb, Rgs6, Pde10a, Zbtb20, Hspa8, Ubash3b, Brinp1, Rimbp2, Chgb, Cdh10, Parp8, Hsp90b1, Mllt3, A330102I10Rik, Pcsk1n, Sparcl1, Map1b, Camk2a, Calr, Phf20l1, Cyp7b1, Clstn2, Prickle1, Kcnq3, Rnf112, Dlgap1, Trim9, Nrn1, Acvr1c, Syn3, Gabrb1, Fbxl7, Mical2, Arpp21, Pcsk2, Sntb2, Kirrel3, Gabbr2, Cdh9, Grm7, Anks1b, Prkag2, Nrg3, Csmd2, Dnajc1, Hectd2, Trpc4, Pdzd2, Mast4, Gm20642, Lrrtm3, Zeb2, Slc35f3, Sez6l, Trim2, Grin2a, Nr3c2, Tspan5, Maml2, Slc2a13, Cacna1c, Hsp90ab1, Itpr1, Atp2b1, Kcnip3, Ksr1, Focad, Tanc1, Arhgap39, Pitpnc1, Auts2, Npy, Acvr1, Dscam, Atp6v0b, Pcdh17, Prkca, Atp6v0a1, Unc79, Sorbs2, Myt1l, Trpm3, Adgrl3, Grm1, Caly, Map2, Sipa1l3, Polb, Cdh2, Tra2a, Cacna2d3, Epha4, Nkain3, Tcf4, Maml3, Pdia6, Plxna4, Hnrnph1, Cdr1os, Ntrk3, Rabgap1l, Rfx3, Slc7a14, Kdm7a, Osbp2, Ank3, Prkce, St18, Grin2b, Cnot3, Car10, Nell2, Hivep2, Ube2e2, Sorl1, Sh3kbp1, Nedd4l, Fam135b, Zmiz1, Cpt1c, Fstl4, Top1, Cdc42bpa, Doc2b, Il1rap, Vps13c, Syt1, Prdm5, Pard3, Ryr2, Sorcs3, Dock4, Aplp2, Negr1, Lrpap1, Rbfox3, Rnf19a, Ptpra, Dscaml1, Tmeff2, Chl1, Klhl2, St7, Dock10, Lgi1, Mycbp2, Olfml2b, Gse1, Nhsl2, Ext1, Pde4d, Neurod2, Dpf3, Nfasc, Pdia3, Cacna1e, Ttc3, Pex5l, Erc2, Pabpc4, Adcy8, Nol4, Chsy3, Etv1, Snhg11, Manf, Rtn4rl1, Galnt17, Rasl10a, Epha5, Akap6, Glt8d1, Slc25a27, Rnasek, Sipa1l1, Bach2, Zbtb16, Penk, Prpf4b, Kcnj6, Kcnq1ot1, Rapgef4, Zdhhc14, Ssh2, A330023F24Rik, Ptprj, Adgrb3, Nudcd3, Adcy9, Plekha5, Specc1, Cspp1, C2cd5, Ptpre, Dlgap2, Slc4a4, Rufy3, Ppp1r9a, Prrt1, Hsp90aa1, Aff2, Lrp1b, Plcl1, Map6, Lrrc7, Ncam2, Gm34567, Inf2, Sema6d, Amph, Unc5d, Nfia, Ctnnd2, Foxp1, Krit1, Hsph1, Zfp608, Efr3a, Nebl, Fgfr1, Hectd2os, Spock3, Nsf, Tbc1d5, Arl6ip1, A930015D03Rik, Dpp6, Zfp804a, Gabra2, Xpr1,Cblb, Chd9, Ptk2b, Nfib, Glis3, Snhg14, Phip, Cbfa2t3, Large1, Stim2, Ptchd4, Ttll11, Gm6999, Herc2, Sorbs1, Fam155a, E130308A19Rik, Mpdz, Slc4a7, Dip2b, Wipf3, Klhl29, Slc24a3, Phf21a, mt-Atp6, Hpca, Carmil1, Igf1r, Plcb1, Dapk1, Dtna, Magi2, Bdnf, Cnih2, Camk1d, Pde4a, Tmem150c, Pgbd5, Dst, Gria1, Aff3, Tle4, Chst1, Ptk2, mt-Co2, Mical3, Bsn, Zdhhc23, Psme4, Meg3, Sorcs2, Sphkap, Slc44a1, Trio, Cnih3, Xbp1, Prkar1b, Aopep, Plekhg5, Tenm1, Tmem108, Gprasp1, Iqgap2, Smoc2, Nlk, Nrgn, Eif4g3, Clvs1, Eml6, Srpk2, Cep112, Dcc, Limch1, Arsj, Erc1, Zfp207, Pam, Mmp17, Lrrc3b, Ppm1h, Frmd4b, Slc25a3, Sybu, Plxna2, Gnal, Sgcz, Ptprn2, Cers4, Cpe, Enox1, Gria2, Usp29, Kcnc1, Adamts17, mt-Co3, Zfp536, Olfm1, Chst11, P4ha1, Sel1l, Arl6ip5, Sema5a, Osbpl6, Foxn3, Dalrd3, Gm44257, Snph, Grm2, B4galt1, Mir99ahg, Nmnat2, Mitf, Slit1, Son, Cst3, Sbf1, Lurap1l, Zmynd8, Npdc1, Mrtfa, Cacnb2, Ncor1, Dgkg, Map2k4, Thy1, Cnpy2, Aig1, Arhgap12, Rph3al, Atrnl1, Emsy, Klhl3, Sipa1l2, Celf2, Rasgrp1, Nktr, Cacna1a, C1ql2, Tanc2, Arhgef12, Gm36975, Plekha2, Crebbp, Pcdh19, Atp6v0c, Dbn1, Gm40841, Ylpm1, Sec62, Ap2a2, Rbm28, Abcb10, Scn3b, Camta1, Exoc4, Celf4, R3hdm1, Pde7b, Selenok, Arid1b, Lingo2, Fnip1, Crlf1, Dok6, Xkr6, Rtn4r, Phf20, Sema3e, Kcnc3, Dot1l, Bcl11b, Ubb, Reep3, Syndig1, Tnfrsf25, Serinc2, Pcnx, Ube2e3, Rere, Kcnma1, Kif21a, Jmjd1c, Ktn1, Bmerb1, Sdccag8, Fkbp2, Ago2, Kcnq2, Tecr, Adgrb1, Csrnp3, Syn2, Dmxl1, Sfpq, Ttc4, Setbp1, Tia1, Coro2b, Dock9, Flnb, Scg2, Spon1, Bbx, Gm12940, Riok3, Rptor, Ccdc88a, Rmdn1, Abhd6, Sdk1, Pacsin2, Mbtd1, Ephb1, Clasp2, Pi4ka, Zfp692, Kcnip2, Mtif2, Fbln2, Syne1, Grik4, Dock11, Gpatch8, Rbfox2, Bcl6, Baiap2, Ramp1, Zc3h13, Lmo4, Dennd1b, Ube3c, Gfod1, Cap2, mt-Cytb, Arl15, Podxl2, Bcl11a, Iqsec3, Cyfip2, Slc25a42, Klhdc10, Cmip, Rian, Kidins220, Anxa11, Ntrk2, Sox11, Ap2b1, Comt, Stk3, Ablim3, Atp1b1, Enah, Slc24a2, Prox1, Spag9)

EE_top100 <- Cs(Nrp1, AC149090.1, Pcdh7, Nrxn3, Tnr, C1ql3, Chst9, Malat1, Hspa5, Dgkb, Rgs6, Pde10a, Zbtb20, Hspa8, Ubash3b, Brinp1, Rimbp2, Chgb, Cdh10, Parp8, Hsp90b1, Mllt3, A330102I10Rik, Pcsk1n, Sparcl1, Map1b, Camk2a, Calr, Phf20l1, Cyp7b1, Clstn2, Prickle1, Kcnq3, Rnf112, Dlgap1, Trim9, Nrn1, Acvr1c, Syn3, Gabrb1, Fbxl7, Mical2, Arpp21, Pcsk2, Sntb2, Kirrel3, Gabbr2, Cdh9, Grm7, Anks1b, Prkag2, Nrg3, Csmd2, Dnajc1, Hectd2, Trpc4, Pdzd2, Mast4, Gm20642, Lrrtm3, Zeb2, Slc35f3, Sez6l, Trim2, Grin2a, Nr3c2, Tspan5, Maml2, Slc2a13, Cacna1c, Hsp90ab1, Itpr1, Atp2b1, Kcnip3, Ksr1, Focad, Tanc1, Arhgap39, Pitpnc1, Auts2, Npy, Acvr1, Dscam, Atp6v0b, Pcdh17, Prkca, Atp6v0a1, Unc79, Sorbs2, Myt1l, Trpm3, Adgrl3, Grm1, Caly, Map2, Sipa1l3, Polb, Cdh2, Tra2a, Cacna2d3)

sobject <- AddModuleScore(sobject,
                  features = list(EE_top500),
                  name= "EE_top500_")

sobject <- AddModuleScore(sobject,
                  features = list(EE_top100),
                  name= "EE_top100_")
```


Module scores - Dot Plot
```{r}
# module scores
modules = Cs(EE_top100_1)

dot.scale = 10 # max size of the dot
scale.min = 0 #lower limit of % expressed
scale.max = 100 #upper limit of % expressed
col.min = -1 #lower limit of avg expression
col.max = 1 #upper limit of avg expression

Idents(sobject) = "condition"
DotPlot(sobject, features = modules,
        col.min = col.min, col.max = col.max, dot.scale = dot.scale, scale.min = scale.min, scale.max = scale.max, dot.min = 0.0) + 
    theme(axis.text.x = element_text(angle = 40, vjust = 1.1, hjust=1.1, size = 18), #axis.text.x = element_text(angle = 0, vjust = 0.2, hjust=0.5, size = 18
          axis.text.y = element_text(size = 18),
          aspect.ratio=1.8, # sets square plot
          axis.line = element_blank(), # hides axes and ticks
          panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
          xlab("") + ylab("") + #coord_flip() 
          #scale_colour_gradient2(low = "#000CFF", mid = "#FFFFFF", high = "#FF0000") +
          scale_colour_gradient2(low = "#FFFFFF", mid = "#FFA7A7", high = "#FF0000") 

Idents(sobject) = "Mouse_ID"
DotPlot(sobject, features = modules,
        col.min = 0, col.max = 1,
        dot.min = 0.0) + 
    theme(axis.text.x = element_text(angle = 0, vjust = 0.2, hjust=0.5, size = 18),
          axis.text.y = element_text(size = 18),
          aspect.ratio=1.2, # sets square plot
          axis.line = element_blank(), # hides axes and ticks
          panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
          xlab("") + ylab("") + #coord_flip() 
          scale_colour_gradient2(high = "brown1") 
```

Get values from DotPlot
```{r}
Idents(sobject) = "Mouse_ID"

p = DotPlot(sobject, features = modules)
p$data

# write to excel file
p_data <- p$data
file_path = "/Users/phinguyen/Downloads"

#library(openxlsx)
write.xlsx(p_data, file = file_path)
```

Module scores - Feature Plot
```{r}
Idents(sobject) = "condition"
label = F
repel = F
label.size = 5
pt.size = 3.5
raster = F
na_cutoff = 0.15
max.cutoff = .5
na_color = "grey75"

features = Cs(RAG_1)

DimPlot(sobject)

FeaturePlot_scCustom(sobject, features = features,
        label = label, repel = repel, label.size = label.size, pt.size = pt.size, order = T, na_cutoff = na_cutoff, max.cutoff=max.cutoff,
        label.color = c("black"), na_color = na_color,
        raster = raster) +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1")

```

Module scores - Violin Plots
```{r}
Idents(sobject) = "condition"
levels(sobject)
pt.size = 0.1

feature = "RAG_1"

VlnPlot(sobject, feature = feature, 
        pt.size = pt.size, slot = 'data') + 
    xlab("")

VlnPlot(sobject, feature = feature, 
        pt.size = pt.size, slot = 'scale.data') + 
    xlab("")
```

